Multiscale Modeling of the Paradoxical Roles of the Immune System in EMT-Mediated Cancer

### 2019-01-11 Paper/Paper.tex
edits up to Results S2

\setcounter{secnumdepth}{4}

\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titlespacing*{\paragraph}
{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
%%

@@

\section{Abstract}
%% Basic Introduction (1-2 sentences)
During tumorigenesis and throughout the lifetime of a tumor, the immune system is constantly engaged in complex interactions in and around the tumor.
%% More Detailed Background (2-3 sentences)
These interactions are influenced by the tumor microenvironment (TME) and in particular by the cellular process of epithelial-to-mesenchymal transition (EMT) in an as-yet unexplained manner.
In turn, both the tumor and the immune system exert substantial influence over the TME in competition to influence the fate of the host.
%% General Problem (1 sentence)
Determining the critical factors and mechanisms involved remains a crucial task for cancer immunotherapy.
%% Summarizing Main Result (1 sentence)
Here, we seek to understand how interplay between the tumor, the immune system, and EMT can determine the outcome of cancer.
%% The Main Result (2-3 sentences)
We developed a multiscale agent-based model that describes tumor-immune interactions and how EMT modulates these.
Exploration of the model revealed that increasing mesenchymal cells' quiescence in some circumstances results in a more tumor-friendly environment and sometimes less.
In contrast, enhancing mesenchymal cells' immune evasiveness always leads to a more tumor-friendly microenvironment.
Moreover, encouraging the TME towards a pro-EMT state also benefits the tumor.
%% Putting into a More General Context (1-2 sentences)
Together, these results indicate that the joint regulation of the TME by the immune system and the tumor itself provides possible targets for future cancer therapies, in particular cancers such as pancreatic cancer in which EMT is speculated to play a role.

@@

\section{Introduction}
Immunotherapy is revolutionizing cancer therapy.
Year-by-year, more breakthroughs in the field produce effective new drugs and therapies that are having a significant impact on patient health and survival.
This is an exciting and fast-paced time in the field and many believe it just the beginning.
There are myriad untapped avenues by which scientists can further explore the role of the immune system in regulating and combatting cancer.
While the direct cell-cell interactions between immune cells and cancer cells is an unsurprising line of inquiry, how immune cells interact with and modulate the tumor microenvironment (TME) is also a promising lead.

This is because immune cells are best understood as one constituent of the TME: responding to the environment's inflammatory signals, fighting the mutated cells present, and then restoring homeostasis in that local context.
Understanding each step in this immune cascade in the context of the TME will be paramount in fully exploring and exploiting all options for immunotherapy.

One aspect of the TME long suspected of playing a vital role in tumor initiation, progression, and metastasis is the epithelial-to-mesenchymal transition (EMT). 
This is the process by which rigidly bound epithelial cells leave a layer of cells and gain motility as mesenchymal cells, among other properties.
As mesenchymal cells often express stem-like properties, their existence has been tied to the initiation and progression of cancer.
Also, due to their motility, some have viewed them as a natural precursor to metastasis.

Thus, herein we explore the interconnectedness of these three features: the tumor itself, the immune system, and the EMT process.
In addition to the interactions already mentioned above, the immune system also affects and is affected by EMT.
The immune system affects EMT by releasing a flurry of cytokines upon arrival in the TME.
In particular, regulatory T cells (Tregs) release copious quantities of the EMT-driving transforming growth factor beta (TGF-$\beta$).
Conversely, the EMTed cells in the tumor downregulate certain surface markers making it more difficult for cytotoxic immune cells to form the immune synapse necessary to lyse the cancer cell.

Taking these three aspects together, all influencing the others, we can consider the simple diagram in Figure \ref{fig:3PartFig} and see just how much potential there is for exploring all these relationships.
Each of these three pieces will be briefly explored in turn for how they will influence one another.

\subsection{The Tumor}\label{TheTumor}
The tumor influences the immune system by its presence and by the cellular processes it engages in that result in shaping the TME.
On the one hand, the mutated DNA of tumor cells results in these cells being recognized as foreign by members of the innate immune system such as Natural Killer cells and Dendritic cells.
It is then these cells that initiate the cascade of cytokines that result in the full activation of the body's immune system to fight the non-self cells.
On the other hand, the tumor cells actively work to shape the immune response that the body launches.
This is accomplished through its own set of cytokines, such as TGF-$\beta$, that work to create a tumor-supportive environment, including shifting the immune balance to be immunosuppressive by enhancing the recruitment of Treg cells.

As mentioned above, TGF-$\beta$ is one driver of EMT, and so it is through direct release of TGF-$\beta$ and recruitment of TGF-$\beta$-secreting Treg cells that the tumor encourages itself and the surrounding tissue towards a more mesenchymal profile.

\subsection{The Immune System}\label{TheImmuneSystem}
The immune system has several competing interactions with the tumor.
While the cytotoxic immune cells, NKs and CTLs, seek out and lyse tumor cells, the regulatory branch of the immune system, Tregs, slows down the effector functions of these other cells.

@@

\subsection{EMT}\label{EMT}
There are many characteristics distinctive of mesenchymal cells and still much research being done in further quantifying these and determining others.
Of interest here, the literature shows two qualities of mesenchymal cells that are of interest particularly in the context of cancer and the immune system.
The first is that mesenchymal cells are less susceptible to immune clearance.
The second is that mesenchymal cells proliferate at a slower rate.

The research on mesenchymal cells' immune evasiveness is quite conclusive with the research even going so far as establishing functional causes for this phenomenon.
Once cytotoxic immune cells, for example natural killer cells and cytotoxic T cells, have targeted a cell to lyse, they necessarily must make a physical connection with that cell, termed an immunological synapse.
However, due to down-regulated surface markers on mesenchymal cells, they are refractory to this immunological task thus rendering them less susceptible to immune clearance.
In line with known motility of mesenchymal cells, scientists have referred to this condition as mesenchymal cells being ``slippery.''
This quality of mesenchymal cells is then quantified in our model by the parameter, MIE, for mesenchymal immune evasion.

The other facet of mesenchymal cells of importance in this research is their slower proliferation rate, and this is tied to the observation that mesenchymal cells often express a more stem-like phenotype.
Indeed, even in healthy tissue, this is readily observed and seen as a useful mechanism for homeostasis as epithelial cells can transition to mesenchymal, stem-like cells under appropriate conditions.
While many qualities are associated with being stem-like, we choose to focus on a reduced proliferation rate.
In quantifying this property, we use the model parameter, MQ, for mesenchymal quiescence.
This choice is less about associating this property with quiescence and more about creating a simple name that easily, if not roughly, captures the essence of what is being modeled.
For example, while mesenchymal proliferation reduction might be a more accurate term, thinking in terms of a negative formulation adds one additional twist to accurately interpreting results which we felt was more detrimental than the possible confusion of invoking ``quiescence.''

@@

\section{Methods}
To address the questions around the relationships among cancer, the immune system, and EMT, we built a multiscale, agent-based model to explore.
The agents in the model are the tissue cells that can either be healthy or have mutations.
It is worth noting that unlike many agent-based models, the spatial arrangement of these cells is not part of the model.
Thus, it is assumed that all the cells are in effect well-mixed as are all the relevant cytokines and other molecules acting on the system.
The immune cells are present in aggregate, i.e. only the total quantity of them present in the tumor microenvironment (TME) is used.
Finally, it is the tissue cells that can be labeled either epithelial or mesenchymal, and this labeling is fluid and depends both on the TME and the internal workings of each individual cell.

@@

Add differential equations? % Adam suggested this, but I'm not sure what this means.
\subsection{The Immune System}\label{ImmuneSystem}
The immune system is modeled as having three active immune cell types: Natural Killer cells (NKs), cytotoxic T cells (CTLs), and T regulatory cells (Tregs).
The NKs and CTLs act on the system by recognizing mutated cells and clearing them.
Should they clear a cell, they themselves are then deactivated and removed from the immune population.
Tregs act on the system by suppressing the effector functions of NKs and CTLs, making it less likely they can clear mutated cells.
In addition to this, Tregs release TGF-$\beta$ which further shapes the TME by pushing tissues cells more towards a mesenchymal phenotype.

@@

\subsection{Running the Model}

\subsubsection{Initializing the Model}
Every model run starts with a fixed amount of cells, $N_0$.
However, different parameter values will lead to different steady states for the population of tissue cells.
In light of this, a warmup period of ~1000 cell cycles is used in which time no mutations are allowed to happen.
Thus, the only immune cell present is NKs.
After the warmup cycles are complete, a mutagenic event is simulated in which all the tissue cells experience a discrete, random jump in their probability to mutate.
From then on, a cell which proliferates either mutates or increases its probability of mutating later on.

\subsubsection{Cell Fate}
During each cell cycle, every cell randomly is assigned a cell fate from the following options:
\begin{itemize}
\item proliferation
\item apoptosis
\item immune clearance (by NKs or CTLs)
\item quiescence
\end{itemize}

@@

\paragraph{Proliferation}
There are four factors that contribute to a cell's probability to weight.
The first is a base proliferation weight that all cells have, $p$.
Second, if the cell has a mutation in the proliferation pathway, then the weight for proliferation is proportionally increased by $\text{prolif}_{\text{up}}$.
Third, if the cell is mesenchymal, then the weight for proliferation is proportionally decreased by $m_Q$, which stands for mesenchymal quiescence.
Fourth, there is a negative feedback of the cells on their own proliferation which is quantified by a Hill factor.
In total, the weight for proliferation is given by

@@

Also, all proliferating cells have a probability of undergoing a pathway mutation.
If they do, one is randomly chosen among the three and that one either mutates or remains mutated.
If they do not mutate, then their probability of mutating increases.

@@

\subsubsection{Cancer Onset}
At the end of each cell cycle, the proportion of tissue cells that are mutated is checked, and if it is above a certain threshold the patient is determined to have developed cancer and a Time to Cancer is recorded.
If not, the process continues until cancer is detected or a preset number of cycles has been completed.
It is also possible with the right combination of parameters, to have patients run out of cells entirely.
The model does not distinguish between these patients and those that have a cancer diagnosis.
The reason is that these patients have succumb to some combination of inflammatory and mutagenic events.

@@

\begin{figure}[H]
\center
\frame{\includegraphics[width=\columnwidth]{Figure1/Figure1.pdf}}
\caption{A. A cartoon of the cell cycle with the immune system interactions as an inset.
B. Global sensitivity analysis.
C. Example survival curve for one cohort.
D. Example of one simulated patient's dynamics. The warmup period for the model is plotted in red.}
\label{fig:ModelIntro}
\end{figure}

\subsection{Exploring the Model}\label{ExplModel}
Refer to Figure \ref{fig:ModelIntro}A for a cartoon of the model.
Here, each cell is represented as a column with three components representing the three possible mutations a cell can have.
If the component is red and has an `X' in the middle, then that indicates the cell has undergone a pathway mutation corresponding to the example cell on the right.
If any component of a cell has a pathway mutation, that cell is then counted as a mutated cell and contributes towards the total number of cancer cells.
Each cell also has a color along the EMT spectrum seen on the left of the cell block. 
Biologically, cells can be anywhere along this continuum with research indicating there are several equilibria along the spectrum.
In our model, this is actually just a binary and that is why the cells shown are at one extreme or the other of the spectrum.

@@

For a typical {\it in silico} patient, see Figure \ref{fig:ModelIntro}B.
Notice that at cell cycle number 541 the proportion of mutated cells reaches 50\% and so the patient has a Time to Cancer of 541.
Once the Time to Cancer has been determined for a patient, most simulations shown later stop.
We do this because once such a high proportion of cells have mutated, the mutant cells quickly take over and make up 100\% of all cells.
Also of interest, as the patient's cells are all rapidly mutating right around this time, most of them are also transitioning to a mesenchymal phenotype with all of them being in a mesenchymal phenotype for a short time during the time period that cancer onset occurs.

Considering the immune populations, notice that the NK population basically stays constant while the lymphocyte populations quickly dwarf the NK population once cells begin to mutate.
These lymphocyte populations exhibit oscillatory behavior but this is an artifact of the inflammation scheme this patient is undergoing.
Every 30 cell cycles the inflammation changes from High to Low or vice versa.
In these roughly month-long periods, the lymphocyte populations increase or decrease rapidly in accordance with this inflammation.

>>>>

\setcounter{secnumdepth}{4}
\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titlespacing*{\paragraph}
{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
%%

@@

\section*{Abstract}
%% Basic Introduction (1-2 sentences)
During tumorigenesis and throughout the lifetime of a tumor, the immune system is constantly engaged in complex interactions in and around the tumor.
%% More Detailed Background (2-3 sentences)
These interactions are influenced by the tumor microenvironment (TME) and in particular by the cellular process of epithelial-to-mesenchymal transition (EMT) in an as-yet unexplained manner.
In turn, both the tumor and the immune system exert substantial influence over the TME in competition to influence the fate of the host.
%% General Problem (1 sentence)
Determining the critical factors and mechanisms involved remains a crucial task for cancer immunotherapy.
%% Summarizing Main Result (1 sentence)
Here, we seek to understand how interplay between the tumor, the immune system, and EMT can determine the outcome of cancer.
%% The Main Result (2-3 sentences)
We developed a multiscale agent-based model that describes tumor-immune interactions and how EMT modulates these.
Exploration of the model revealed that increasing mesenchymal cells' quiescence in some circumstances results in a more tumor-friendly environment and sometimes less.
In contrast, enhancing mesenchymal cells' immune evasiveness always leads to a more tumor-friendly microenvironment.
Moreover, encouraging the TME towards a pro-EMT state also benefits the tumor.
%% Putting into a More General Context (1-2 sentences)
Together, these results indicate that the joint regulation of the TME by the immune system and the tumor itself provides possible targets for future cancer therapies, in particular cancers such as pancreatic cancer in which EMT is speculated to play a role.

@@

\section{Introduction}
Immunotherapy is slowly revolutionizing how we treat cancer.
Year-by-year, more breakthroughs in the field produce effective new drugs and therapies that are having a significant impact on patient health and survival. Ref  [examples] 
There are myriad untapped avenues by which scientists can further explore the role of the immune system in regulating and combatting cancer; in addition to direct cell-cell interactions between immune and cancer cells, the means by which immune cells interact with and modulate the tumor microenvironment (TME) is an important research area. Immune cells can be understood as one constituent of the TME: responding to the environment's inflammatory signals, fighting the mutated cells present, and then restoring homeostasis in that local context. Ref 
Understanding each step in this immune cascade in the context of the TME will be paramount in fully exploring and exploiting all options for immunotherapy.

Recent work on cancer-immune interactions [briefly describe Guo paper]...

Epithelial-to-mesenchymal transition (EMT) describes the (reversible) cell plasticity process by which epithelial cells transition into mesenchymal cells. EMT has an important yet complicated role in tumor initiation, progression, and metastasis. Ref 
As mesenchymal cells often exhibit stem-like properties, their existence has been tied to the initiation and progression of cancer. Ref
Also, due to their motility, some have viewed them as a natural precursor to metastasis. Ref

Herein, we explore the interconnectedness of these three key systems: the tumor, the immune system, and EMT.
In addition to the interactions already discussed, the immune system affects and is affected by EMT.
The immune system affects EMT by releasing a flurry of cytokines upon arrival in the TME. Ref 
In particular, regulatory T cells (Tregs) release copious quantities of the EMT-driving transforming growth factor beta (TGF-$\beta$). Ref 
Conversely, the EMTed cells in the tumor downregulate certain surface markers making it more difficult for cytotoxic immune cells to form the immune synapse necessary to lyse the cancer cell.
Taking these three aspects together, all influencing the others, we can consider the simple diagram in Figure \ref{fig:3PartFig} and see just how much potential there is for exploring all these relationships.
Each of these three pieces will be briefly explored in turn for how they will influence one another.

[No subsections headings within Introduction]
\subsection{The Tumor}\label{TheTumor}
The tumor influences the immune system by its presence and by the cellular processes it engages in that result in shaping the TME.
On the one hand, the mutated DNA of tumor cells results in these cells being recognized as foreign by members of the innate immune system such as natural killer (NK) cells and dendritic cells.
It is then these cells that initiate the cascade of cytokines that result in the full activation of the body's immune system to fight the non-self cells.
On the other hand, the tumor cells actively work to shape the immune response that the body launches.
This is accomplished through its own set of cytokines, such as transforming growth factor beta (TGF-$\beta$), that work to create a tumor-supportive environment, including shifting the immune balance to be immunosuppressive by enhancing the recruitment of T regulatory (Treg) cells. Since TGF-$\beta$ is also a driver of EMT, then through direct release of TGF-$\beta$ and recruitment of TGF-$\beta$-secreting Treg cells, cells of the tumor and the surrounding tissue are driven towards a more mesenchymal profile.

\subsection{The Immune System}\label{TheImmuneSystem}
The immune system has several competing interactions with the tumor.
While the cytotoxic immune cells, NKs and [define!] CTLs, seek out and lyse tumor cells, the regulatory branch of the immune system, Tregs, slows down the effector functions of these other cells.

@@

\subsection{EMT}\label{EMT}
There are many characteristics distinctive of mesenchymal cells and still much research being done in further quantifying these and determining others.
Of interest here, the literature shows two qualities of mesenchymal cells that are of interest particularly in the context of cancer and the immune system.
The first is that mesenchymal cells are less susceptible to immune clearance. Ref 
The second is that mesenchymal cells proliferate at a slower rate. Ref

The research on mesenchymal cells' immune evasiveness is quite conclusive with the research even going so far as establishing functional causes for this phenomenon.
Once cytotoxic immune cells, for example natural killer cells and cytotoxic T cells, have targeted a cell to lyse, they necessarily must make a physical connection with that cell, termed an immunological synapse. Ref 
However, due to down-regulated surface markers on mesenchymal cells, they are refractory to this immunological task thus rendering them less susceptible to immune clearance. Ref 
In line with known motility of mesenchymal cells, scientists have referred to this condition as mesenchymal cells being ``slippery.''
This quality of mesenchymal cells is then quantified in our model by the parameter, MIE, for mesenchymal immune evasion.

The other facet of mesenchymal cells of importance in this research is their slower proliferation rate, and this is tied to the observation that mesenchymal cells often express a more stem-like phenotype.
Indeed, even in healthy tissue, this is readily observed and seen as a useful mechanism for homeostasis as epithelial cells can transition to mesenchymal, stem-like cells under appropriate conditions.
While many qualities are associated with being stem-like, we choose to focus on a reduced proliferation rate.
In quantifying this property, we use the model parameter `mesenchymal growth arrest', MGA.
This choice is less about associating this property with quiescence and more about creating a simple name that easily, if not roughly, captures the essence of what is being modeled.
For example, while mesenchymal proliferation reduction might be a more accurate term, thinking in terms of a negative formulation adds one additional twist to accurately interpreting results which we felt was more detrimental than the possible confusion of invoking ``quiescence.''

In the following sections of the paper we first introduce the model modularly, which is appropriate diven its multiscale and hybrid nature. We then perform sensitivity analysis and identify parameters that are crucial for the progression to tumorigenesis. We analyze the competing effects of EMT and of the immune system on tumorigenesis and find that EMT intricately regulates tumorigenesis: under certain regimes a careful balance of EMT- and immune-driven process can significantly prolong cancer-free survival.

@@

\section{Methods}
[In this section it is important to say what is old (from previous model), i.e. `following the model of XX, we....' and what is new/our addition.]

To address the questions around the relationships among cancer, the immune system, and EMT, we built a multiscale, agent-based model.
The agents in the model are the tissue cells with tumorigenic potential, that can either be mutation-free, or have mutations.
It is worth noting that we do not consider spatial dependencies between agents, i.e. we assume that cells are well-mixed as are all the relevant cytokines and other molecules acting on the system.
We model the effects of immune cell types and cytokines as continuous fields within the tumor microenvironment (TME), since here single-cell resolution is not required. 
Finally, tissue cells that can be labeled either epithelial or mesenchymal, and this labeling is fluid and depends both on the TME and the internal workings of each individual cell.

@@

Add differential equations? % Adam suggested this, but I'm not sure what this means.
% I think this comment applied to the next section: add the equations describing immune system

\subsection{The Immune System}\label{ImmuneSystem}
The immune system is modeled as having three active immune cell types: Natural Killer cells (NKs), cytotoxic T cells (CTLs), and T regulatory cells (Tregs).
The NKs and CTLs act on the system by recognizing mutated cells and clearing them.
Should they clear a cell, they themselves are then deactivated and removed from the immune population.
Tregs act on the system by suppressing the effector functions of NKs and CTLs, making it less likely they can clear mutated cells.
In addition to this, Tregs release TGF-$\beta$ which further shapes the TME by pushing tissues cells more towards a mesenchymal phenotype.

@@

\subsection{Model Simulation}

\subsubsection{Initial conditions} Every model run starts with a fixed amount of cells, $N_0$.
However, different parameter values will lead to different steady states for the population of tissue cells.
In light of this, a warmup period of ~1000 cell cycles is used in which time no mutations are allowed to happen.
Thus, the only immune cell present is NKs.
After the warmup cycles are complete, a mutagenic event is simulated in which all the tissue cells experience a discrete, random jump in their probability to mutate.
From then on, a cell which proliferates either mutates or increases its probability of mutating later on.

\subsubsection{Tissue cell fate}
During each cell cycle, every cell randomly is assigned a cell fate from the following options:
\begin{itemize}
\item proliferation
\item apoptosis
\item immune clearance (by NKs or CTLs)
\item quiescence
\end{itemize}

@@

\paragraph{Proliferation}
There are four factors that contribute to a cell's probability of proliferation.
% NB try to avoid posessive apostrophe. I.e. replace this with
There are four factors that contribute to the probability that a cell will proliferate.
% 
The first is a base proliferation probability that all cells have, $p$.
Second, if the cell has a mutation in the proliferation pathway, then the weight for proliferation is proportionally increased by $\text{prolif}_{\text{up}}$.
Third, if the cell is mesenchymal, then the weight for proliferation is proportionally decreased by $m_Q$, which stands for mesenchymal quiescence.
Fourth, there is a negative feedback of the cells on their own proliferation which is quantified by a Hill factor.
In total, the weight for proliferation is given by

@@

Also, all proliferating cells have a probability of undergoing a pathway mutation.
If they do, one is randomly chosen among the three and that one either mutates or remains mutated.
If they do not mutate, then their probability of mutating increases. [??]

@@

\subsubsection{Progression and tumorigenesis}
At the end of each cell cycle, the proportion of tissue cells that are mutated is calculated, and if it is above a certain threshold, a tumorigenic event is recorded and the patient is determined to have cancer (using a similar strategy to Guo). The Time to Cancer is given as the time from the start of the simulation until tumorigenesis. Simulations run until either a tumorigenic event is recorded or until the maximum number of cycles has been reached.
%It is also possible with the right combination of parameters, to have patients run out of cells entirely.
%The model does not distinguish between these patients and those that have a cancer diagnosis.
%The reason is that these patients have succumb to some combination of inflammatory and mutagenic events.

@@

\begin{figure}[H]
\center
\frame{\includegraphics[width=\columnwidth]{Figure1/Figure1.pdf}}
\caption{A. Schematic depiction of the model with blue/red denoting non-mutated/mutated cells in the agent-based model. In each cycle, the fate of a cell is regulated by the EMT state of the cell and by any pathway mutations harbored. Inset shows major processes of immune system regulation on the tissue cells.
B. Example simulation of the dynamics of a one patient; warmup period is shown in red.
C. Example survival curve for one cohort of patients with parameters XXXXX.
D. Global sensitivity analysis of XXX parameters using XXXXX.}
\label{fig:ModelIntro}
\end{figure}

\subsection{Exploring the Model}\label{ExplModel}
In Figure \ref{fig:ModelIntro}A a schematic description of the model is shown.
Here, each cell is represented as a column with three components representing the three possible mutations a cell can have.
If the component is red and has an `X' in the middle, then that indicates the cell has undergone a pathway mutation corresponding to the annotation to the right of the cell block.
If any component [???] of a cell has a pathway mutation, that cell is then counted as a mutated cell and contributes towards the total number of cancer cells.
The EMT score of each cell (denoting position along the EMT spectrum) is associated with a color given by the legend to the left of the cell block. 
In the model the EMT spectrum is continuous. While biologically, there may be only a few discrete stable cell states available on the EMT spectrum, recent studies highlight a growing number of states accessible, with research indicating there are several equilibria along the spectrum. Ref [Tian and Levine lab] 
In our model, this is actually just a binary and that is why the cells shown are at one extreme or the other of the spectrum.

@@

For a typical {\it in silico} patient, see Figure \ref{fig:ModelIntro}B.
Notice that at cell cycle number 541 the proportion of mutated cells reaches 50\% and so the patient has a Time to Cancer of 541. [translate into days?]
Once the Time to Cancer has been determined for a patient, most simulations shown later stop. [??? mean to say: we stop the simulation?]
We do this because once such a high proportion of cells have mutated, the mutant cells quickly take over and make up 100\% of all cells.
Also of interest, as the patient's cells are all rapidly mutating right around this time, most of them are also transitioning to a mesenchymal phenotype with all of them being in a mesenchymal phenotype for a short time during the time period that cancer onset occurs.

Considering the immune populations, notice that the NK population basically stays constant while the lymphocyte populations quickly dwarf the NK population once cells begin to mutate.
These lymphocyte populations exhibit oscillatory behavior, however this is not due to intrinsic dynamics but rather due to the inflammation scheme that this patient is undergoing: every 30 cell cycles, the inflammation level changes from high to low, or vice versa. In each of these periods periods, the lymphocyte populations increase or decrease rapidly in accordance with the inflammation.
